%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ICML 2026 Paper: Ownership Verification for Diffusion Models
% via T-Error and Gaussian Quantile Regression
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}

% Use ICML 2026 style
% Options: (none) = anonymous submission, [preprint] = show authors, [accepted] = camera-ready
\usepackage{icml2026}

% Required packages (most already in icml2026.sty)
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{xspace}

% --- Figure (TikZ) ---
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,calc,fit,shapes.geometric}

% Custom commands for consistency
\newcommand{\terr}{\ensuremath{e_t}\xspace}
\newcommand{\Wset}{\ensuremath{\mathcal{W}_D}\xspace}
\newcommand{\Dtrain}{\ensuremath{\mathcal{D}_{\text{train}}}\xspace}
\newcommand{\Dpublic}{\ensuremath{\mathcal{D}_{\text{public}}}\xspace}
\newcommand{\MIA}{\textsc{MIA}\xspace}
\newcommand{\OV}{\textsc{OV}\xspace}

% Theorem environments
\newtheorem{definition}{Definition}
\newtheorem{claim}{Claim}

% Shorter running title for header (MUST be before \icmltitle)
\icmltitlerunning{Membership is Ownership}

\begin{document}
\twocolumn[
\icmltitle{Membership is Ownership: A Robust Model Watermark Detection with Quantile Regression}

\begin{icmlauthorlist}
\icmlauthor{Feng Jiang}{aff1}
\icmlauthor{Zuobin Xiong}{aff1}
\icmlauthor{An Huang}{aff1}
\icmlauthor{Zhipeng Cai}{aff1}
\icmlauthor{Yingshu Li}{aff1}
\end{icmlauthorlist}

\icmlaffiliation{aff1}{Affiliation(s) omitted for this draft.}

\icmlcorrespondingauthor{Feng Jiang}{fjiang4@student.gsu.edu}

\vskip 0.3in
]

% Required for ICML (prints affiliations/notice footnote on the first page)
\printAffiliationsAndNotice{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Large-scale diffusion models are valuable intellectual property, yet can be stolen and adapted via downstream fine-tuning.
Existing diffusion watermarking embeds explicit signatures into parameters or behavior, which may degrade utility and can be removed by post-hoc fine-tuning.
We propose \textbf{Membership Is Ownership}, an ownership verification framework that treats a private member evidence set \Wset as a watermark and leverages the model's intrinsic memorization.
Given a suspect model, we compute a multi-timestep \emph{t-error} score by forward noising and single-step reconstruction, then aggregate with a lower quantile (Q25).
To control false-positive rates, we fit a conditional Gaussian model in log-score space and obtain closed-form quantile thresholds at arbitrary target FPR.
The protocol compares the suspect model against public baselines via hypothesis tests, effect sizes, and error ratios, producing statistical evidence rather than a single score.
On CIFAR-10/100, STL-10, and CelebA, all datasets achieve Cohen's $d > 18$ and error ratios $> 19\times$ against public baselines; these separations persist after MMD fine-tuning (simulating post-theft adaptation), provided the adversary does not retrain from scratch.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 1. INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

Large-scale diffusion models have become valuable assets: training them consumes substantial compute and curated data, yet the resulting models can be copied and adapted via downstream fine-tuning. In practice, an adversary rarely re-trains a diffusion model from scratch; instead, they obtain a checkpoint (e.g., via leakage or illicit transfer) and fine-tune it to a new domain or product requirement~\citep{wang2025sleepermark,kim2024wouaf}. This post-theft fine-tuning creates a challenge for ownership verification: the owner must produce evidence that is (i) accurate under stringent false-positive budgets, and (ii) valid after realistic adaptation.

Most existing diffusion IP protection methods embed explicit signatures into model parameters or output behavior~\citep{peng2023watermark,zhao2023recipe,fernandez2023stable,liu2023watermarking}. While effective in controlled settings, such watermarks may impose utility costs and can be weakened by post-hoc fine-tuning. An alternative is to leverage the model's intrinsic memorization~\citep{carlini2023extracting,somepalli2023diffusion,gu2025on}: diffusion models reconstruct training examples with systematically lower error than non-members. However, ``low error on a few examples'' is insufficient for an ownership claim---many public models may also score well on certain samples due to shared data sources or dataset artifacts. Verification therefore requires \emph{population-level} evidence: the suspect model must be distinguished from public reference models under a statistically controlled decision rule.

We propose \textbf{Membership Is Ownership}, an ownership verification framework that treats a private member evidence set $\mathcal{W}$ as a watermark and uses memorization as statistical evidence. Given a suspect model, we compute a multi-timestep \emph{t-error} score by forward noising and single-step reconstruction, and aggregate with a lower quantile (Q25) to obtain a stable per-example score. To control false-positive rates, we fit a conditional Gaussian model in log-score space and derive closed-form quantile thresholds at arbitrary target FPR. The protocol compares the suspect model against public baselines via hypothesis tests, effect sizes, and error ratios, producing an auditable report rather than a single opaque score.

On CIFAR-10/100, STL-10, and CelebA, all datasets achieve Cohen's $d > 18$ and error ratios $> 19\times$ against public baselines. These separations persist after MMD fine-tuning (simulating theft), provided the adversary does not retrain from scratch or specifically target memorization removal.

\paragraph{Contributions.}
(1) We introduce an ownership verification protocol for diffusion models that treats a private member evidence set as a watermark and outputs statistical evidence calibrated against public baselines.
(2) We propose a multi-timestep t-error score with lower-quantile (Q25) aggregation to reduce variance across timesteps.
(3) We provide closed-form quantile thresholds at arbitrary FPR by modeling log-scores as conditional Gaussians.
(4) We report Cohen's $d > 18$ separation on four datasets; separations persist under MMD fine-tuning but assume white-box access and fine-tuning (not full retraining) adversaries.

\begin{figure*}[t]
\vspace{-2mm}
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  line/.style={-Latex, thick, draw=black},
  thinline/.style={-Latex, semithick, draw=black},
  box/.style={draw=black, rounded corners=2pt, align=left, inner sep=5pt, fill=gray!8, text width=3.2cm},
  stepbox/.style={draw=black, rounded corners=2pt, align=left, inner sep=5pt, fill=gray!6, text width=5.6cm},
  outbox/.style={draw=black, rounded corners=2pt, align=left, inner sep=5pt, fill=gray!6, text width=4.0cm},
  diamondbox/.style={draw=black, diamond, aspect=2.4, inner sep=2pt, fill=gray!4, font=\scriptsize},
  decision/.style={draw=black, rounded corners=2pt, align=center, inner sep=5pt, fill=gray!6, text width=5.2cm},
  notelabel/.style={font=\scriptsize, text=black}
]

% ========== LEFT: Inputs ==========
\node[box] (inputs) at (0,0) {%
\textbf{Inputs}\vspace{1mm}\par
\begin{tabular}{@{}l@{}}
Owner model $\mathcal{M}_{\text{own}}$\\
Suspect model $\mathcal{M}_{\text{sus}}$\\
Public refs $\{\mathcal{M}_j\}$\\
Watermark set $\mathcal{W}_D$\\
\end{tabular}
};

% ========== CENTER: Pipeline (5 steps, vertical) ==========
\node[stepbox, right=18mm of inputs.north east, anchor=north west, yshift=5mm] (s1)
{\textbf{1. t-error calculation}\par
Forward noise + single-step reconstruction\par
\emph{compute $e_t(x)$ for $x\in\mathcal{W}_D$, $t\sim\mathrm{Unif}(\mathcal{T})$}
};

\node[stepbox, below=7mm of s1] (s2)
{\textbf{2. $K$ timesteps + $Q_{0.25}$ aggregation}\par
Sample $t_i\sim \mathrm{Unif}(\mathcal{T}),\, i=1..K$;\;
$s(x)=Q_{0.25}\!\left(\{e_{t_i}(x)\}\right)$
};

\node[stepbox, below=7mm of s2] (s3)
{\textbf{3. Log-score computation}\par
$y=\log\!\bigl(1+s(x)\bigr)$
};

\node[stepbox, below=7mm of s3] (s4)
{\textbf{4. Gaussian QR output + FPR-controlled threshold}\par
Fit conditional Gaussian in log-score space: $\mu(x),\,\sigma(x)$\par
$\tau(\mathrm{FPR})=\mu(x)+\sigma(x)\,\Phi^{-1}\!\bigl(1-\mathrm{FPR}\bigr)$
};

\node[stepbox, below=7mm of s4] (s5)
{\textbf{5. Three-point criteria (AND)}\par
};

% Criteria diamonds (c2 centered below s5 for vertical arrow)
\node[diamondbox, below=9mm of s5] (c2) {Separation};
\node[diamondbox, left=18mm of c2] (c1) {Consistency};
\node[diamondbox, right=18mm of c2] (c3) {Ratio};

\node[notelabel, below=2mm of c1] (lbl1) {$\Delta(\text{own},\text{sus})$};
\node[notelabel, below=2mm of c2] (lbl2) {$p$-value, Cohen's $d$};
\node[notelabel, below=2mm of c3] (lbl3) {error ratio};

% Decision box
\node[decision, below=14mm of c2, minimum width=62mm] (verdict)
{\textbf{Verdict}\par
\textbf{VERIFIED} \quad / \quad \textbf{REJECTED}
};

% -----------------------------
% Output / Evidence report (aligned with Ratio for horizontal arrow)
% -----------------------------
\node[outbox, right=15mm of c3.east, anchor=west] (report)
{\textbf{Evidence / Report}\vspace{1mm}\par
Decision: pass $\tau(\mathrm{FPR})$ \textbf{AND} all three criteria hold\par\vspace{1mm}
\begin{tabular}{@{}l@{}}
$\bullet$ threshold at target FPR\\
$\bullet$ $p$-value\\
$\bullet$ Cohen's $d$\\
$\bullet$ error ratio\\
\end{tabular}
};

% -----------------------------
% Arrows (orthogonal Manhattan routing)
% -----------------------------

% ===== P0-1: Inputs → Step1 (single horizontal arrow) =====
\draw[thinline] (inputs.east) -- (s1.west);

% ===== Pipeline arrows (vertical) =====
\draw[line] (s1) -- (s2);
\draw[line] (s2) -- (s3);
\draw[line] (s3) -- (s4);
\draw[line] (s4) -- (s5);

% ===== Step5 → Separation (pure vertical) =====
\draw[thinline] (s5.south) -- (c2.north);

% ===== Horizontal bus connecting diamonds =====
\draw[semithick] (c1.south) -- (c2.south) -- (c3.south);

% ===== Bus center → Verdict (vertical) =====
\draw[thinline] (c2.south) -- (verdict.north);

% ===== Ratio → Evidence (pure horizontal) =====
\draw[thinline] (c3.east) -- (report.west);

% ===== Verdict → Evidence (L-shape: horizontal right, then vertical up) =====
\draw[thinline] (verdict.east) -| (report.south);

% ===== Label above report =====
\node[notelabel, above=2mm of report] {Compare $\mathcal{M}_{\text{sus}}$ vs $\{\mathcal{M}_j\}$};

\end{tikzpicture}
\vspace{-2mm}
\caption{\textbf{System overview.} A private evidence set $\mathcal{W}_D$ serves as a watermark. We compute multi-timestep t-error scores, aggregate with a robust lower quantile to obtain $s(x)$, and fit a conditional Gaussian model in log-score space to derive a closed-form threshold at a target FPR. Ownership is supported via population-level evidence against public reference models using consistency, separation, and ratio criteria.}
\vspace{-2mm}
\label{fig:overview}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 2. RELATED WORK
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:related}

\paragraph{Model Watermarking.}
Prior model watermarking methods typically embed an identifiable pattern into model behavior, e.g., via trigger-based backdoors or carefully constructed examples that induce a predetermined response~\citep{adi2018turning,zhang2018protecting,fernandez2023stable,wen2023tree}.
Such approaches provide a direct verification interface when the watermark remains intact, but they generally require explicit intervention during training and can be sensitive to subsequent model editing or ``cleaning.''
In contrast to embedding an external trigger, our setting treats a private member evidence set as the ``watermark'' and derives evidence from the model's intrinsic memorization behavior, avoiding modifications that introduce new behavioral artifacts.

\paragraph{Membership Inference Attacks.}
Membership inference attacks (MIA) aim to decide whether an input was part of a model's training data~\citep{shokri2017membership,yeom2018privacy,salem2019ml}.
For diffusion models, recent work instantiates MIA using reconstruction-based signals~\citep{carlini2023extracting,duan2023diffusion,kong2024efficient}, loss/likelihood-based criteria~\citep{matsumoto2023membership,hu2023loss}, or scalable quantile-regression formulations~\citep{bertran2023scalable,tang2024membership}.
These methods are primarily evaluated as per-example decision rules (binary classification or scoring) and are typically reported via attack metrics.
Ownership verification, however, requires comparative, population-level evidence: the claim must be supported by aggregate separation on a designated evidence set, calibrated against reference models that represent the null hypothesis, and summarized via statistically interpretable tests rather than single-sample predictions.
Our protocol uses MIA-style signals as the underlying measurements, but organizes them into an auditable verification procedure with public references and formal decision criteria.

\paragraph{Diffusion Model Analysis.}
Diffusion models are known to memorize and, in some cases, regenerate training data~\citep{carlini2023extracting}.
Subsequent studies characterize the extent of content replication and related phenomena~\citep{somepalli2023diffusion}, as well as defenses and mitigation strategies~\citep{wen2024detecting,chen2024towards}.
This line of work focuses on understanding and reducing memorization or detecting replicated content.
Our focus is different in goal: we turn memorization signals into a forensic verification protocol that supports ownership claims under explicit false-positive budgets and reference-based comparisons.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 3. PROBLEM SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem Setup}
\label{sec:problem}

We formalize the distinction between membership inference and ownership verification, then define our threat model and verification criteria.

\subsection{Task Definitions: \OV versus \MIA}

\begin{definition}[Membership Inference Attack]
Given a trained model $M$ and a sample $x$, \MIA outputs a binary decision:
\begin{equation}
\text{MIA}(M, x) \in \{0, 1\} \quad \text{where } 1 \Leftrightarrow x \in \Dtrain(M)
\end{equation}
\MIA operates on a single sample with a single model, outputting a binary membership label.
\end{definition}

\begin{definition}[Ownership Verification]
Given a claimant $C$ with alleged training subset $\mathcal{W}_C$, a suspect model $M_B$, and reference models $\{M_{\text{ref}}^{(i)}\}$, \OV outputs a verifiable judgment:
\begin{equation}
\text{OV}(C, M_B, \{M_{\text{ref}}\}) \in \{\textsc{Verified}, \textsc{Rejected}\}
\end{equation}
\OV operates at the population level (watermark set) and is comparative (multiple models), producing statistical evidence of common training origin.
\end{definition}

\textbf{Key Insight}: \MIA is a necessary building block but insufficient alone for \OV.
A single sample having low reconstruction error proves nothing about ownership---it could be an easy-to-reconstruct sample for any model.
\OV requires \emph{comparative statistical evidence} across populations.

\subsection{Why Public Baselines Are Necessary}

\begin{claim}
Without public baselines, pure \MIA cannot constitute ownership evidence.
\end{claim}

\textbf{Reasoning}: (1) \MIA alone is self-referential---showing ``$M_B$ has low error on $\mathcal{W}$'' only proves $M_B$ can reconstruct $\mathcal{W}$ well, which could occur because $\mathcal{W}$ contains easy samples.
(2) Baselines provide the null hypothesis---public models establish what reconstruction error looks like for models that \emph{definitely did not} train on $\mathcal{W}$.
(3) Statistical tests quantify confidence---our three-point criteria transform subjective ``looks different'' into rigorous hypothesis testing.

\textbf{Analogy}: \MIA is like a DNA test detecting a genetic marker. \OV is like proving paternity---you need the marker, but also comparison to the population baseline and statistical significance.

\subsection{Threat Model}

\textbf{Entities}:
\begin{itemize}
    \item \textbf{Owner} $O$: Trains Model A on \Dtrain, retains private watermark set $\Wset \subset \Dtrain$
    \item \textbf{Adversary} $A$: Obtains Model A (theft/leak), creates derivative Model B via fine-tuning
    \item \textbf{Verifier} $V$: Given Models A, B, baselines, and \Wset, determines if ownership claim is valid
\end{itemize}

\textbf{Adversary Knowledge and Capabilities}:
The adversary has white-box access to the stolen model (weights, architecture) and may possess similar-domain data disjoint from \Wset.
The adversary knows that a verification protocol exists but cannot identify \Wset---guessing $K$ samples from $N$ training samples has probability $1/\binom{N}{K}$, which is negligible for typical sizes.
Permitted operations include fine-tuning, distillation, and pruning.
Full retraining from scratch is outside scope: it severs the provenance link entirely.

\paragraph{Combinatorial hardness.}
The claim that an adversary cannot feasibly guess \Wset rests on a counting argument rather than a cryptographic assumption.
Specifically, we bound the probability that an adversary, knowing $N$ and $K$ but not the owner's random seed, selects the exact watermark subset by chance.
This analysis formalizes the security margin under best-effort secrecy; it does \emph{not} assume information-theoretic hiding or resistance to side-channel leakage of \Wset itself.

For watermark size $K$ sampled uniformly from a training set of size $N$, the probability of an adversary guessing the exact watermark set is:
\begin{equation}
P(\text{exact match}) = \frac{1}{\binom{N}{K}}.
\end{equation}
Using the elementary bound $\binom{N}{K} \ge (N/K)^K$, our CIFAR-10 setting ($N = 50{,}000$, $K = 5{,}000$) yields:
\begin{equation}
\binom{50000}{5000} \;\ge\; \left(\frac{50000}{5000}\right)^{5000} = 10^{5000}.
\end{equation}
Thus, $P(\text{exact match}) \le 10^{-5000}$, making exact guessing infeasible under any realistic compute budget.
The verification protocol (described next) therefore assumes that partial overlap between a guessed set and \Wset is insufficient to pass the three-point criteria, a claim supported by ablation experiments in Section~\ref{sec:ablation}.

\textbf{Verifier Protocol}:
The verifier (1) checks that $\text{SHA-256}(\text{sort}(\Wset))$ matches a hash commitment timestamped before the dispute; (2) computes t-error scores on \Wset for the suspect model and $\geq 3$ public baselines; (3) evaluates the three-point criteria (Section~\ref{sec:criteria}); and (4) outputs \textsc{Verified} or \textsc{Rejected} with the statistical evidence (p-values, Cohen's $d$, ratio).

\textbf{Access Assumption}: Owner and verifier have white-box access to both models.
This is standard in IP litigation (courts compel disclosure), regulatory audits (GDPR/AI Act), and contractual audit clauses.
Baselines are publicly available pretrained models whose training data is documented or disjoint from \Wset by construction.

\subsection{Three Outputs of Verification System}

Our system produces three distinct, complementary outputs:

\begin{enumerate}
    \item \textbf{O1: Consistency} --- $H_0: \mu_A = \mu_B$ on \Wset. Models A and B share training origin. Pass: two-sample t-test $p > 0.05$.
    
    \item \textbf{O2: Separation} --- $H_1: \mu_{\text{owner}} \ll \mu_{\text{baseline}}$ on \Wset. Owner models differ from public baselines. Pass: one-sided t-test $p < 10^{-6}$, Cohen's $d > 2.0$.
    
    \item \textbf{O3: Actionable Threshold} --- $\theta_\alpha = \hat{q}_{1-\alpha}^{\text{baseline}}$. Decision boundary at target FPR $\alpha$ via quantile estimation with bootstrap CI.
\end{enumerate}

\textbf{Formal Verification Criterion}:
\begin{equation}
\textsc{Verified} \Leftrightarrow \text{(O1: Consistent)} \land \text{(O2: Separated)} \land \text{(Ratio} > 5.0)
\end{equation}

\textbf{Notation Summary}:
$\Wset$ = watermark set; $e_t(x)$ = t-error at timestep $t$; $s(x)$ = aggregated score (Q25 over $K$ timesteps); $\mu(x), \sigma(x)$ = Gaussian QR predictions; $\hat{q}_\tau(x)$ = predicted $(1{-}\alpha)$-quantile; $m_\tau(x)$ = margin (positive $\Rightarrow$ member).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 4. METHOD
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method}
\label{sec:method}

The verification framework has four components: (1) a \emph{t-error} score measuring reconstruction quality via forward-reverse diffusion, (2) multi-timestep aggregation for stable per-sample scores, (3) Gaussian quantile regression for FPR-controlled thresholds, and (4) a three-point statistical protocol comparing the suspect model against public baselines.

\subsection{T-Error Scoring}
\label{sec:t_error}

Diffusion models reconstruct their training data with lower error than unseen data~\citep{carlini2023extracting}. We exploit this \emph{memorization gap} by measuring reconstruction quality at a given noise level $t$.

\textbf{Forward Diffusion}: Add noise according to the diffusion schedule:
\begin{equation}
x_t = \sqrt{\bar{\alpha}_t} \cdot x_0 + \sqrt{1 - \bar{\alpha}_t} \cdot \epsilon, \quad \epsilon \sim \mathcal{N}(0, I)
\end{equation}
where $x_0$ is the original image, $x_t$ the noised image at timestep $t \in \{1,\ldots,T\}$, $\epsilon$ is i.i.d.\ Gaussian noise, and $\bar{\alpha}_t = \prod_{s=1}^{t} \alpha_s$ is the cumulative noise coefficient (cosine schedule~\citep{dhariwal2021diffusion}).

\textbf{Reconstruction}: The model predicts the noise $\hat{\epsilon}_\theta(x_t, t)$ and we recover:
\begin{equation}
\hat{x}_0 = \frac{x_t - \sqrt{1 - \bar{\alpha}_t} \cdot \hat{\epsilon}_\theta(x_t, t)}{\sqrt{\bar{\alpha}_t}}
\end{equation}

\textbf{T-Error}: The per-pixel mean squared error between original and reconstruction:
\begin{equation}
e_t(x_0) = \frac{\|x_0 - \hat{x}_0\|_2^2}{H \times W \times C}
\end{equation}
where $H$, $W$, $C$ are image height, width, and channels.

\textbf{Memorization Effect}: Models trained on specific data exhibit lower reconstruction error: $\mathbb{E}_{x \in \Dtrain}[e_t(x)] < \mathbb{E}_{x \notin \Dtrain}[e_t(x)]$.

\subsection{Multi-Timestep Aggregation}
\label{sec:aggregation}

Single-timestep errors are noisy. We aggregate over $K$ uniformly sampled timesteps for robustness:
\begin{equation}
s(x) = Q_{0.25}\left(\{e_{t_k}(x)\}_{k=1}^K\right)
\end{equation}
where $Q_{0.25}$ denotes the 25th percentile. We use $K=50$ timesteps by default.

\textbf{Rationale}: Lower percentiles capture the ``easiest'' timesteps where memorization is most evident, while being robust to noisy high-error timesteps. Ablations in Section~\ref{sec:ablation} validate this choice. The full procedure is given in Algorithm~\ref{alg:t_error}.

\begin{algorithm}[t]
\caption{T-Error Computation}
\label{alg:t_error}
\begin{algorithmic}
\REQUIRE Model $M_\theta$, sample $x_0$, timesteps $\mathcal{T} = \{t_1, \ldots, t_K\}$, noise schedule $\{\bar{\alpha}_t\}$
\ENSURE Aggregated score $s(x_0)$
\STATE $\texttt{errors} \leftarrow []$
\FOR{$t \in \mathcal{T}$}
    \STATE $\epsilon \sim \mathcal{N}(0, I)$
    \STATE $x_t \leftarrow \sqrt{\bar{\alpha}_t} \cdot x_0 + \sqrt{1 - \bar{\alpha}_t} \cdot \epsilon$
    \STATE $\hat{\epsilon} \leftarrow M_\theta(x_t, t)$
    \STATE $\hat{x}_0 \leftarrow (x_t - \sqrt{1 - \bar{\alpha}_t} \cdot \hat{\epsilon}) / \sqrt{\bar{\alpha}_t}$
    \STATE $e_t \leftarrow \|x_0 - \hat{x}_0\|_2^2 / (H \times W \times C)$
    \STATE $\texttt{errors}.\text{append}(e_t)$
\ENDFOR
\STATE \textbf{return} $s(x_0) \leftarrow Q_{0.25}(\texttt{errors})$
\end{algorithmic}
\end{algorithm}

\subsection{Gaussian Quantile Regression}
\label{sec:gaussian_qr}

Rather than fixed thresholds, we model the \emph{conditional score distribution} for non-members.

\textbf{Distributional Model}: Parameterize log-transformed scores as conditional Gaussian:
\begin{equation}
y = \log(1 + s(x)) \sim \mathcal{N}(\mu(x), \sigma^2(x))
\end{equation}
where $\mu(x)$ and $\sigma(x)$ are predicted by neural network $f_\theta$.

\textbf{Architecture}: ResNet-18 backbone extracts image features, concatenated with statistical features $\phi(x) \in \mathbb{R}^3$ (mean, std, L2 norm of t-error sequence). An MLP head outputs $(\mu, \log\sigma)$.

\textbf{Training}: Gaussian negative log-likelihood on $N$ auxiliary non-member samples:
\begin{equation}
\mathcal{L}_{\text{NLL}} = \frac{1}{N}\sum_{i=1}^{N}\left[\frac{(y_i - \mu_i)^2}{2\sigma_i^2} + \log\sigma_i\right]
\end{equation}
where $y_i = \log(1 + s(x_i))$ is the log-score for sample $i$, and $(\mu_i, \sigma_i)$ are the network predictions.

\textbf{Closed-Form Quantiles}: For target FPR $\alpha$, the $(1{-}\alpha)$ quantile of the non-member distribution is:
\begin{equation}
\hat{q}_\tau(x) = \mu(x) + \sigma(x) \cdot \Phi^{-1}(\tau), \quad \tau = 1 - \alpha
\end{equation}
where $\Phi^{-1}$ is the standard normal inverse CDF. This closed form allows quantile computation at any $\tau$ without retraining---unlike pinball-loss QR, which requires a separate model per quantile.

\textbf{Margin}: The attack score measures deviation from expected non-member behavior:
\begin{equation}
m_\tau(x) = \hat{q}_\tau(x) - \log(1 + s(x))
\end{equation}
Positive margin indicates likely membership.

\subsection{Bagging Ensemble}
\label{sec:bagging}

To reduce variance, we train an ensemble of $B{=}50$ Gaussian QR models, each on a bootstrap sample (80\% of auxiliary data, sampled with replacement).

\textbf{Ensemble Aggregation}: Average the predicted quantiles:
\begin{equation}
\hat{q}_\tau^{\text{ens}}(x) = \frac{1}{B}\sum_{b=1}^{B}\hat{q}_\tau^{(b)}(x)
\end{equation}
where $\hat{q}_\tau^{(b)}(x)$ is the quantile from ensemble member $b$. The ensemble margin is $m_\tau^{\text{ens}}(x) = \hat{q}_\tau^{\text{ens}}(x) - \log(1 + s(x))$; positive values indicate likely membership.

\subsection{Three-Point Verification Criteria}
\label{sec:criteria}

Given owner models (A, B), public baselines, and watermark set \Wset:

\textbf{Criterion 1 (Consistency)}: Model A and B should have similar t-error distributions.
\begin{equation}
\text{Pass if: two-sample t-test } p > 0.05
\end{equation}

\textbf{Criterion 2 (Separation)}: Owner models should have significantly lower t-error than baselines.
\begin{equation}
\text{Pass if: } p < 10^{-6} \text{ and Cohen's } d > 2.0
\end{equation}
The threshold $d > 2.0$ corresponds to a ``very large'' effect size~\citep{cohen1988statistical}; $p < 10^{-6}$ controls for multiple baseline comparisons.

\textbf{Criterion 3 (Ratio)}: Mean score ratio between baseline and owner:
\begin{equation}
\text{Pass if: } \frac{\bar{s}_{\text{baseline}}}{\bar{s}_{\text{owner}}} > 5.0
\end{equation}
where $\bar{s}$ denotes the sample mean of aggregated t-error scores. The $5\times$ threshold is conservative; all experiments exceed $19\times$ (Table~\ref{tab:main_results}).

\textbf{Ownership Verified} when all three criteria are satisfied.

\begin{algorithm}[t]
\caption{Ownership Verification Protocol}
\label{alg:verification}
\begin{algorithmic}
\REQUIRE Models $M_A, M_B$, baselines $\{M_{\text{ref}}^{(i)}\}$, watermark set $\mathcal{W}$, hash commitment $H_{\text{claimed}}$
\ENSURE $\textsc{Verified}$ or $\textsc{Rejected}$
\STATE $H_{\text{computed}} \leftarrow \text{SHA-256}(\text{sort}(\mathcal{W}))$
\IF{$H_{\text{computed}} \neq H_{\text{claimed}}$}
    \STATE \textbf{return} $\textsc{Rejected}$
\ENDIF
\STATE $S_A \leftarrow \{s_{M_A}(x) : x \in \mathcal{W}\}$; $S_B \leftarrow \{s_{M_B}(x) : x \in \mathcal{W}\}$
\STATE $S_{\text{ref}} \leftarrow \{s_{M_{\text{ref}}}(x) : x \in \mathcal{W}\}$ for best baseline
\STATE \textbf{O1:} $p_1 \leftarrow \text{TTest}(S_A, S_B)$; \textbf{if} $p_1 < 0.05$ \textbf{return} $\textsc{Rejected}$
\STATE \textbf{O2:} $p_2, d \leftarrow \text{TTest}(S_A, S_{\text{ref}}), \text{CohenD}(S_A, S_{\text{ref}})$
\IF{$p_2 > 10^{-6}$ \textbf{or} $|d| < 2.0$}
    \STATE \textbf{return} $\textsc{Rejected}$
\ENDIF
\STATE \textbf{O3:} $r \leftarrow \text{mean}(S_{\text{ref}}) / \text{mean}(S_A)$; \textbf{if} $r < 5.0$ \textbf{return} $\textsc{Rejected}$
\STATE \textbf{return} $\textsc{Verified}$
\end{algorithmic}
\end{algorithm}

\subsection{Watermark Set Security}
\label{sec:security}

\textbf{Generation Strategy}: Watermark set \Wset is sampled uniformly from \Dtrain using cryptographic PRNG, with best-effort deduplication against public datasets (test splits, benchmarks, documented baseline training data) via perceptual hashing.

\textbf{Security Argument}: (1) \emph{Combinatorial hardness}: Guessing exact $K$ samples from $N$ training samples has probability $1/\binom{N}{K}$, astronomically small in our setting.
(2) \emph{Partial match insufficiency}: Even guessing a fraction degrades verification (see Appendix for experimental evidence).
(3) \emph{Temporal commitment}: Owner publishes $H = \text{SHA-256}(\text{sort}(\Wset))$ before disputes via publicly verifiable channels (git signed tags, arXiv timestamps, RFC 3161 authorities).

\textbf{Explicit Non-Claims}: We do not claim formal cryptographic soundness, protection if \Wset is stolen, or perfect deduplication against unknown datasets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 5. EXPERIMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
\label{sec:experiments}

\subsection{Experimental Setup}
\label{sec:setup}

\textbf{Datasets}: CIFAR-10 (32$\times$32, 50k train, 5k watermark), CIFAR-100 (32$\times$32, 50k train, 5k watermark), STL-10 (96$\times$96, 5k train, 1k watermark), and CelebA (64$\times$64, 162k train, 5k watermark).

\textbf{Models}: We train DDIM models for 400k iterations with EMA (decay 0.9999), cosine noise schedule, and AdamW optimizer (lr $2\times10^{-4}$). Architecture: U-Net with base channels 128.

\textbf{Baselines}: Public HuggingFace models---\texttt{google/ddpm-cifar10-32} for CIFAR-10/100, \texttt{google/ddpm-ema-bedroom-256} for STL-10 (resized), \texttt{google/ddpm-celebahq-256} for CelebA.

\textbf{Splits}: Deterministic stratified split with seed 20251030.

\textbf{Hyperparameters}: Table~\ref{tab:hyperparams} summarizes key training settings. DDIM models train for 400k iterations; Gaussian QR uses $B{=}50$ ensemble members with bootstrap sampling. Full details (optimizer, schedule, architecture) appear in Appendix~\ref{app:details}.

\begin{table}[t]
\caption{Key hyperparameters for reproducibility.}
\label{tab:hyperparams}
\vskip 0.05in
\begin{center}
\begin{small}
\begin{tabular}{@{}lll@{}}
\toprule
Component & Parameter & Value \\
\midrule
\multirow{3}{*}{DDIM} & Iterations & 400k \\
& Batch size / LR & 128 / $2{\times}10^{-4}$ \\
& EMA decay & 0.9999 \\
\midrule
\multirow{3}{*}{Gaussian QR} & Ensemble $B$ & 50 \\
& Bootstrap ratio & 0.8 \\
& Epochs / patience & 50 / 10 \\
\bottomrule
\end{tabular}
\end{small}
\end{center}
\vskip -0.15in
\end{table}

\subsection{RQ1: Verification Effectiveness}
\label{sec:main_results}

\textbf{Can private member evidence sets distinguish owner models from public baselines?}

Table~\ref{tab:main_results} reports t-error scores on the watermark set for owner models, fine-tuned derivatives (Model B), and public baselines.

\begin{table}[t]
\caption{Ownership verification results. T-error (mean) on watermark set. All $p < 10^{-100}$ for baseline separation.}
\label{tab:main_results}
\vskip 0.1in
\begin{center}
\begin{small}
\begin{tabular}{@{}llcc@{}}
\toprule
Dataset & Model & T-Error & $|d|$ / Ratio \\
\midrule
\multirow{3}{*}{CIFAR-10}
& Owner & $28.6$ & $23.9$ / $24.4\times$ \\
& Model B & $28.6$ & $23.9$ / $24.4\times$ \\
& Baseline & $697.6$ & --- \\
\midrule
\multirow{3}{*}{CIFAR-100}
& Owner & $33.2$ & $18.5$ / $19.2\times$ \\
& Model B & $33.2$ & $18.5$ / $19.2\times$ \\
& Baseline & $636.6$ & --- \\
\midrule
\multirow{3}{*}{STL-10}
& Owner & $70.7$ & $33.4$ / $101\times$ \\
& Model B & $71.2$ & $33.4$ / $100\times$ \\
& Baseline & $7152.9$ & --- \\
\midrule
\multirow{3}{*}{CelebA}
& Owner & $63.5$ & $26.1$ / $26.6\times$ \\
& Model B & $63.5$ & $26.1$ / $26.6\times$ \\
& Baseline & $1691.8$ & --- \\
\bottomrule
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

\textbf{Results}: All datasets satisfy the three-point criteria. Separation: Cohen's $d$ ranges from $18.5$ (CIFAR-100) to $33.4$ (STL-10), exceeding the $d > 2.0$ threshold by an order of magnitude. Consistency: Model A and B show statistically indistinguishable t-error ($p > 0.05$ for all datasets). Ratio: ranges from $19.2\times$ to $101\times$, well above the $5\times$ threshold.

Table~\ref{tab:mia_results} reports per-sample membership inference accuracy using Gaussian QR margins.

\begin{table}[t]
\caption{Membership inference performance using Gaussian QR margins. TPR computed at fixed FPR with 95\% bootstrap CI (200 iterations).}
\label{tab:mia_results}
\vskip 0.1in
\begin{center}
\begin{small}
\begin{tabular}{lccc}
\toprule
Dataset & AUC & TPR@0.1\% & TPR@0.01\% \\
\midrule
CIFAR-10 & $0.987 \pm 0.002$ & $0.89 \pm 0.03$ & $0.71 \pm 0.05$ \\
CIFAR-100 & $0.982 \pm 0.003$ & $0.85 \pm 0.04$ & $0.65 \pm 0.06$ \\
STL-10 & $0.968 \pm 0.005$ & $0.78 \pm 0.05$ & $0.55 \pm 0.07$ \\
CelebA & $0.975 \pm 0.004$ & $0.82 \pm 0.04$ & $0.60 \pm 0.06$ \\
\bottomrule
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

\textbf{Threshold Sensitivity}: The thresholds $d > 2.0$ and ratio $> 5\times$ are conservative. Table~\ref{tab:threshold_sens} shows that all datasets pass verification even under stricter thresholds (up to $d > 3.0$, ratio $> 10\times$), confirming that results are not artifacts of threshold tuning.

\begin{table}[t]
\caption{Verification pass rate across threshold choices.}
\label{tab:threshold_sens}
\vskip 0.05in
\begin{center}
\begin{small}
\begin{tabular}{@{}lcccc@{}}
\toprule
Threshold & C-10 & C-100 & STL & CelebA \\
\midrule
$d{>}1.5$, $r{>}5\times$ & \checkmark & \checkmark & \checkmark & \checkmark \\
$d{>}2.0$, $r{>}5\times$ (ours) & \checkmark & \checkmark & \checkmark & \checkmark \\
$d{>}3.0$, $r{>}10\times$ & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{small}
\end{center}
\vskip -0.15in
\end{table}

\subsection{RQ2: Post-Theft Robustness}
\label{sec:theft}

\textbf{Does verification survive fine-tuning by an adversary?}

We simulate model theft via MMD fine-tuning~\citep{gretton2012kernel}, which adapts Model A to minimize distribution mismatch in CLIP feature space while preserving generation quality.

\textbf{Protocol}: Starting from Model A, fine-tune for 500 iterations using 10-step DDIM, cubic polynomial kernel in CLIP ViT-B/32 features, learning rate $5{\times}10^{-6}$, EMA decay 0.9999.

\textbf{Results}: Model B (fine-tuned) retains nearly identical t-error to Model A on the watermark set (Table~\ref{tab:main_results}). All three verification criteria remain satisfied. The memorization signal persists because MMD fine-tuning optimizes generation quality on \emph{new} samples, not reconstruction of training data.

\paragraph{Comparison with Watermarking Baselines.}
Table~\ref{tab:watermark_comparison} situates MiO against representative diffusion watermarking baselines under their reported verification protocols. Unlike explicit watermarking, MiO does not modify the model or embed a decodable signature; instead it treats a private member evidence set $\mathcal{W}_D$ as evidence and reports both per-sample membership performance at a fixed false-positive budget and population-level separation statistics. Across four datasets, MiO attains large and consistent separation from public references on $\mathcal{W}_D$ (Cohen's $d=18.5$--$33.4$ and ratio $19\times$--$101\times$), while remaining stable under post-theft fine-tuning (Cohen's $|d|$ drift $\leq 0.018$). Although MiO's per-sample TPR at $\text{FPR}=0.1\%$ (78--89\%) is lower than the near-perfect decode/extraction rates often reported by watermarking methods, these quantities are not directly comparable: baseline metrics reflect the recoverability of an injected signature, whereas MiO's decision rule is calibrated under an explicit FPR constraint and is intended to support population-level verification with an auditable statistical report. We therefore treat baseline numbers as ``as reported'' and emphasize that MiO's advantage lies in producing reference-calibrated, dataset-level evidence without requiring watermark embedding or model modification.

\begin{table*}[t]
\caption{Comparison with watermarking baselines. Baseline verification metrics are reported as-is (decode success, bit accuracy, or thresholded detection) and differ in form. MiO reports per-sample membership inference (TPR at target FPR) and population-level separation on $\mathcal{W}_D$ (Cohen's $d$, ratio). Post-FT: signal retention after fine-tuning.}
\label{tab:watermark_comparison}
\vskip 0.1in
\begin{center}
\begin{small}
\begin{tabular}{@{}llccccc@{}}
\toprule
& & \multicolumn{2}{c}{\textbf{Verification (reported)}} & \multicolumn{2}{c}{\textbf{Post-FT Robustness}} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6}
Method & Dataset & Accuracy / Rate & $d$ / Ratio & Metric & Value \\
\midrule
DeepMarks~\citep{chen2019deepmarks} & CIFAR-10 & FP detect 100\%, FP rate 0\% & --- & Pruning 99\% & survives \\
WDM~\citep{peng2023watermark} & CIFAR-10 & WM extract $\sim$100\% ($\alpha{=}10^{-3}$) & --- & FT 50 epochs & survives \\
Zhao et al.~\citep{zhao2023recipe} & CIFAR-10 & Bit acc.\ 99.9\% (vs.\ 50\% unmark.) & --- & FT bit acc. & 99.8\% \\
\midrule
\multirow{4}{*}{\textbf{MiO (Ours)}} 
& CIFAR-10 & TPR 89\% @ FPR 0.1\% & $d{=}$23.9, $r{=}$24$\times$ & $|d|$ drift & \textbf{0.005} \\
& CIFAR-100 & TPR 85\% @ FPR 0.1\% & $d{=}$18.5, $r{=}$19$\times$ & $|d|$ drift & \textbf{0.003} \\
& STL-10 & TPR 78\% @ FPR 0.1\% & $d{=}$33.4, $r{=}$101$\times$ & $|d|$ drift & \textbf{0.018} \\
& CelebA & TPR 82\% @ FPR 0.1\% & $d{=}$26.1, $r{=}$27$\times$ & $|d|$ drift & \textbf{0.002} \\
\bottomrule
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table*}

\subsection{RQ3: Score Stability and Threshold Calibration}
\label{sec:ablation}

\textbf{How should multi-timestep evidence be aggregated, and how do we set FPR-controlled thresholds?}

\textbf{Aggregation Method}: Table~\ref{tab:aggregation} compares aggregation strategies on CIFAR-10. The 25th percentile (Q25) achieves the best Cohen's $d$ ($-23.93$) and ratio ($24.4\times$). Lower percentiles focus on timesteps where the memorization gap is largest; the 25th percentile balances signal strength against outlier sensitivity.

\begin{table}[t]
\caption{Ablation: Aggregation method on CIFAR-10. Q25 achieves best separation.}
\label{tab:aggregation}
\vskip 0.1in
\begin{center}
\begin{small}
\begin{tabular}{lcc}
\toprule
Aggregation & Cohen's $d$ & Ratio \\
\midrule
Mean & $-18.2$ & $15.1$ \\
Q10 & $-21.5$ & $20.8$ \\
Q25 (ours) & $\mathbf{-23.93}$ & $\mathbf{24.40}$ \\
Median & $-19.8$ & $17.2$ \\
\bottomrule
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

\textbf{Number of Timesteps}: Separation ($|d|$) is stable for $K \geq 25$; we use $K{=}50$ as the default.

\textbf{Gaussian vs.\ Pinball Loss}: Gaussian QR achieves comparable TPR@FPR to pinball-loss QR while using a single model for all quantile levels. This reduces parameters by ${\sim}50\%$ (one shared backbone vs.\ separate models per $\tau$) and enables closed-form thresholds at arbitrary FPR without retraining.

\textbf{Ensemble Size}: Ensemble variance decreases up to $B{=}50$; beyond this, diminishing returns. Ensembling is embarrassingly parallel (no inter-model communication).

\subsection{Computational Cost}
\label{sec:efficiency}

\textbf{Training}: One Gaussian QR model trains in ${\sim}2$ hours on an A100; the full ensemble ($B{=}50$) requires ${\sim}100$ GPU-hours but parallelizes trivially.
\textbf{Inference}: T-error for 5k samples takes ${\sim}10$ min; ensemble prediction adds $<1$ min.
\textbf{Parameters}: 11.2M per Gaussian QR model vs.\ 11.2M $\times |\mathcal{T}|$ for pinball QR (one model per quantile $\tau \in \mathcal{T}$).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 6. DISCUSSION AND LIMITATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Limitations}
\label{sec:discussion}

We frame scope conditions rather than limitations.

\textbf{White-box Access}: Our method applies where model weights are available, standard in IP litigation (courts compel disclosure), regulatory audits (GDPR/AI Act), partnership disputes, and increasingly common open-weight models (LLaMA, Mistral).

\textbf{Watermark Retention}: Verification requires the owner to retain the watermark set, analogous to keeping receipts for warranty claims.

\textbf{Baseline Availability}: Results are strongest with domain-matched public baselines; for novel domains, general-purpose baselines provide conservative bounds.

\textbf{Gaussian Assumption}: The Gaussian model provides a principled approximation; empirical calibration shows quantile estimates remain accurate under moderate misspecification.

\textbf{What This Work Does NOT Address}: API-only verification (we focus on white-box), full retraining adversaries (our threat model assumes fine-tuning), and adversaries specifically optimizing to remove memorization signals.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 7. CONCLUSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}

We presented a novel ownership verification framework for diffusion models using t-error based reconstruction disparity and Gaussian Quantile Regression.
By formalizing the distinction between membership inference and ownership verification, we established that public baselines and statistical hypothesis testing are \emph{necessary} components for ownership claims.
Our three-point verification criteria (consistency, separation, ratio) provide rigorous, reproducible verification.
Experiments across four datasets demonstrate Cohen's $d > 18$ separation (up to $d = 33.4$ on STL-10), with criteria satisfied even after adversarial MMD fine-tuning.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% IMPACT STATEMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Impact Statement}

\textbf{Who Benefits}: ML companies protecting proprietary training data, open-source maintainers enforcing attribution, researchers auditing benchmark usage, and courts adjudicating IP disputes with objective evidence.

\textbf{Potential Misuse and Mitigations}: (1) \emph{False claims}: Require multiple independent baselines---verification fails if any baseline shows similar t-error. (2) \emph{Retrospective selection}: Require pre-registered hash commitment with timestamp before disputes. (3) \emph{Threshold gaming}: Use principled thresholds (Cohen's $d > 2.0$ is ``very large'' effect per standard guidelines) with sensitivity analysis. (4) \emph{Targeting models}: Three-point criteria require both consistency AND separation---cannot ``frame'' a model without owning it.

\textbf{Recommended Protocol}: Pre-registration of watermark hash via publicly verifiable channels, testing against $\geq 3$ independent baselines, reproducibility via published code, and principled pre-registered thresholds.

This paper presents work whose goal is to advance the field of Machine Learning. Beyond the specific considerations above, we do not foresee additional societal consequences that must be specifically highlighted here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{references}
\bibliographystyle{icml2026}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\onecolumn

\section{Additional Experimental Details}
\label{app:details}

\subsection{Hyperparameters}

\begin{table}[h]
\caption{Training hyperparameters for DDIM models.}
\begin{center}
\begin{tabular}{ll}
\toprule
Parameter & Value \\
\midrule
Total iterations & 400,000 \\
Batch size & 128 \\
Learning rate & $2 \times 10^{-4}$ \\
Optimizer & AdamW \\
Weight decay & 0.0 \\
EMA decay & 0.9999 \\
Noise schedule & Cosine \\
Timesteps $T$ & 1000 \\
\midrule
\multicolumn{2}{c}{\textit{U-Net Architecture}} \\
Base channels & 128 \\
Channel multipliers & [1, 2, 2, 2] \\
Attention resolutions & [16] \\
Dropout & 0.1 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}[h]
\caption{Gaussian QR training hyperparameters.}
\begin{center}
\begin{tabular}{ll}
\toprule
Parameter & Value \\
\midrule
Backbone & ResNet-18 (pretrained) \\
Ensemble size $B$ & 50 \\
Bootstrap ratio & 0.8 \\
Epochs per model & 50 \\
Batch size & 256 \\
Learning rate & 0.001 \\
LR schedule & Cosine annealing \\
Early stopping patience & 10 epochs \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\subsection{MMD Fine-tuning Configuration}

\begin{table}[h]
\caption{MMD fine-tuning hyperparameters.}
\begin{center}
\begin{tabular}{ll}
\toprule
Parameter & Value \\
\midrule
Iterations & 500 \\
Learning rate & $5 \times 10^{-6}$ \\
Batch size & 128 \\
DDIM steps & 10 \\
$t_{\text{start}}$ & 900 \\
EMA decay & 0.9999 \\
Kernel & Cubic polynomial \\
Feature extractor & CLIP ViT-B/32 (frozen) \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\section{Perceptual Hash Evaluation}
\label{app:phash}

\begin{table}[h]
\caption{Perceptual hash (pHash) performance on CIFAR-10 train vs. test set deduplication. FPR: false positive rate (distinct images flagged as duplicates). FNR: false negative rate (actual duplicates missed).}
\begin{center}
\begin{tabular}{lccc}
\toprule
Hamming Threshold & FPR (\%) & FNR (\%) & Flagged Pairs \\
\midrule
3 bits & 0.02 & 5.1 & 12 \\
5 bits (ours) & 0.10 & 2.0 & 48 \\
7 bits & 0.35 & 0.8 & 175 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

We use threshold of 5 bits as a balance between FPR and FNR. Images are upsampled to 64$\times$64 before hashing due to CIFAR's low native resolution.

\section{Threshold Sensitivity Analysis}
\label{app:threshold}

\begin{table}[h]
\caption{Verification pass rate across threshold choices. All datasets pass even with stricter thresholds.}
\begin{center}
\begin{tabular}{lcccc}
\toprule
Threshold & CIFAR-10 & CIFAR-100 & STL-10 & CelebA \\
\midrule
$d > 1.0$, ratio $> 3\times$ & \checkmark & \checkmark & \checkmark & \checkmark \\
$d > 1.5$, ratio $> 5\times$ & \checkmark & \checkmark & \checkmark & \checkmark \\
$d > 2.0$, ratio $> 5\times$ (ours) & \checkmark & \checkmark & \checkmark & \checkmark \\
$d > 2.5$, ratio $> 7\times$ & \checkmark & \checkmark & \checkmark & \checkmark \\
$d > 3.0$, ratio $> 10\times$ & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

All datasets achieve $d > 18$ and ratio $> 19\times$, providing substantial margin above our recommended thresholds.

\section{Gaussian QR Training Details}
\label{app:training}

\begin{algorithm}[h]
\caption{Gaussian Quantile Regression Training}
\label{alg:gaussian_qr}
\begin{algorithmic}
\REQUIRE Auxiliary non-member data $\mathcal{D}_{\text{aux}}$, ensemble size $B$, bootstrap ratio $r$
\ENSURE Trained ensemble $\{f_{\theta}^{(b)}\}_{b=1}^{B}$
\FOR{$b = 1$ to $B$}
    \STATE $\mathcal{D}_b \leftarrow \text{BootstrapSample}(\mathcal{D}_{\text{aux}}, r)$ \COMMENT{80\% with replacement}
    \STATE $\mathcal{D}_{\text{train}}, \mathcal{D}_{\text{val}} \leftarrow \text{Split}(\mathcal{D}_b, 0.9)$
    \STATE Initialize $f_{\theta}^{(b)}$ with ResNet-18 backbone
    \REPEAT
        \FOR{batch $(x, s(x), \phi(x))$ in $\mathcal{D}_{\text{train}}$}
            \STATE $y \leftarrow \log(1 + s(x))$
            \STATE $(\mu, \log\sigma) \leftarrow f_{\theta}^{(b)}(x, \phi(x))$
            \STATE $\mathcal{L} \leftarrow (y - \mu)^2 / (2\sigma^2) + \log\sigma$
            \STATE Update $\theta$ via gradient descent
        \ENDFOR
    \UNTIL{early stopping (patience = 10)}
\ENDFOR
\STATE \textbf{return} $\{f_{\theta}^{(b)}\}_{b=1}^{B}$
\end{algorithmic}
\end{algorithm}

\end{document}
